---
title: Descriptive Statistics in R
author: not sure
description: snippets for producing descriptive statistics in R
original dat:
date:


## R provides a wide range of functions for obtaining summary statistics. One method of obtaining descriptive statistics is to use the 
## sapply( ) function with a specified summary statistic. 

# get means for variables in data frame mydata
# excluding missing values
```{r echo=FALSE}
sapply(mydata, mean, na.rm=TRUE) 

```

# Possible functions used in sapply include mean, sd, var, min, max, median, range, and quantile.

# There are also numerous R functions designed to provide a range of descriptive statistics at once. For example

```{r}

# mean,median,25th and 75th quartiles,min,max
summary(mydata)

```

# Tukey min,lower-hinge, median,upper-hinge,max
```{r}
fivenum(x)

```

# Using the Hmisc package
```{r}
library(Hmisc)
describe(mydata)
# n, nmiss, unique, mean, 5,10,25,50,75,90,95th percentiles
# 5 lowest and 5 highest scores
```

# Using the pastecs package
```{r}
library(pastecs)
stat.desc(mydata)
# nbr.val, nbr.null, nbr.na, min max, range, sum,
# median, mean, SE.mean, CI.mean, var, std.dev, coef.var
```

# Using the psych package
```{r}
library(psych)
describe(mydata)
# item name ,item number, nvalid, mean, sd,
# median, mad, min, max, skew, kurtosis, se
```

## Summary Statistics by Group
# A simple way of generating summary statistics by grouping variable is available in the psych package.
```{r}
library(psych)
describe.by(mydata, group,...)
```

# The doBy package provides much of the functionality of SAS PROC SUMMARY. It defines the desired table using a model formula and a function. Here is a simple example.
```{r}
library(doBy)
summaryBy(mpg + wt ~ cyl + vs, data = mtcars,
          FUN = function(x) { c(m = mean(x), s = sd(x)) } )
# produces mpg.m wt.m mpg.s wt.s for each
# combination of the levels of cyl and vs 
```