---
title: R Project Template
author: James W. Vizzard
description: A basic R project checklist to get started so I don't have to go back and unscrew things like Git integration
original date: 20260330
tags: data, r, template
---

# Use the "Github first" option to start the project.
# Use GitKraken to create a new repository. Give it the name to use on both Github and the local machine.
# Copy the repository URL from Github
# In RStudio, create a new project using Version Control > Git
# Enter the repository URL in the popup
# Set the local directory to /Data/*projectname* [project name = Github repository name]
# The Git tab in RStudio should show the repository as an option

```{r echo=FALSE, message=FALSE, error=FALSE}
# Use pacman to load and install dependencies.
library(pacman)

# List your projects's dependencies here, comma separated:
p_load(tidyverse, dplyr, here, readr, haven, knitr, ggplot2, RColorBrewer)

# set base folder for data
# here::i_am("Documents/Qatar_Cars")
```

# load data and label variable columns
```{r echo=FALSE, message=FALSE, error=FALSE}
# load and label the data
# load qatarcars.csv
qcars <- read_csv ("qatarcars.csv")

# label the columns
attr(qcars$origin, "label") <- "Country of Origin"
attr(qcars$make, "label") <- "Manufacturer"
attr(qcars$model, "label") <- "Model"
attr(qcars$length, "label") <- "Length (m)"
attr(qcars$width, "label") <- "Width (m)"
attr(qcars$height, "label") <- "Height (M)"
attr(qcars$seating, "label") <- "Number of seats"
attr(qcars$trunk, "label") <- "Trunk capacity (L)"
attr(qcars$economy, "label") <- "Fuel Economy (L/100km)"
attr(qcars$horsepower, "label") <- "Horsepower"
attr(qcars$price, "label") <- "2025 price ($US)"
attr(qcars$mass, "label") <- "Weight (KG)"
attr(qcars$performance, "label") <- "0-100 KPH"
attr(qcars$type, "label") <- "Body Type"
attr(qcars$enginetype, "label") <- "Fuel"
```

## R provides a wide range of functions for obtaining summary statistics. One 
## method of obtaining descriptive statistics is to use the sapply( ) function 
## with a specified summary statistic. 

# get means for variables in data frame mydata excluding missing values
```{r echo=FALSE}
sapply(mydata, mean, na.rm=TRUE) 

```

# Possible functions used in sapply include mean, sd, var, min, max, median, 
#range, and quantile.

# There are also numerous R functions designed to provide a range of descriptive
# statistics at once. For example

```{r}

# mean,median,25th and 75th quartiles,min,max
summary(mydata)

```

# Tukey min,lower-hinge, median,upper-hinge,max
```{r}
fivenum(x)
```

# Using the Hmisc package
```{r}
library(Hmisc)
describe(mydata)
# n, nmiss, unique, mean, 5,10,25,50,75,90,95th percentiles
# 5 lowest and 5 highest scores
```

# Using the pastecs package
```{r}
library(pastecs)
stat.desc(mydata)
# nbr.val, nbr.null, nbr.na, min max, range, sum,
# median, mean, SE.mean, CI.mean, var, std.dev, coef.var
```

# Using the psych package
```{r}
library(psych)
describe(mydata)
# item name ,item number, nvalid, mean, sd,
# median, mad, min, max, skew, kurtosis, se
```

## Summary Statistics by Group
# A simple way of generating summary statistics by grouping variable is 
# available in the psych package.
```{r}
library(psych)
describe.by(mydata, group,...)
```

# The doBy package provides much of the functionality of SAS PROC SUMMARY. It 
# defines the desired table using a model formula and a function. Here is a 
# simple example.
```{r}
library(doBy)
summaryBy(mpg + wt ~ cyl + vs, data = mtcars,
          FUN = function(x) { c(m = mean(x), s = sd(x)) } )
# produces mpg.m wt.m mpg.s wt.s for each
# combination of the levels of cyl and vs 
```

## Structure the data set
# Create a subset of only gas-powered cars (using Musgrave's Qatar_Cars)
```{r echo=FALSE, message=FALSE, error=FALSE}
petrolcars <- subset( x = qcars,                    # data frame is qatarcars
              subset = enginetype != "Electric"     # subset of petrol and hybrid cars
              )
```

# Create a subset that removes the columns for height, width, and length
```{r echo=FALSE, message=FALSE, error=FALSE}

pcarsmass <- petrolcars %>%
  select(origin, make, model, economy, price, mass)
```

# Select model
```{r}
#reference: https://www.math.ucla.edu/~anderson/rw1001/library/base/html/step.html

petrolcarsmod <- lm(pcarsmass$price ~ pcarsmass$economy + pcarsmass$make + pcarsmass$origin + pcarsmass$mass)
  
#step assess the model to determine the best model--danger of overfitting

bestmodel <- step(petrolcarsmod)
bestmodel
```

## Create plots
# Create scatter plot of horsepower and economy with selected color pallette
```{r fig.width=6, fig.height=4}
# Scatterplot
p <- ggplot(data=petrol, mapping=aes(x = horsepower, y = economy, color = make))
p + geom_point(size = 2) + scale_color_brewer(palette = "Set1") + theme(legend.position = "top")
ggsave(filename = "scatterplot_horsepower_economy.png").  # save plot as .png

# Boxplot
b <- ggplot(data = petrolcars,
            mapping = aes(x = origin, y = economy, color = origin))
b + geom_boxplot() +
labs(x = "Economy", y = "Country",
     title = "Economy of Petrol & Hybrid Cars by Country",
     subtitle = "L/100KM",
     caption = "Source: Paul Musgrave")
ggsave(filename = "boxplot_national_economy.png").       # save plot as .png
```

